Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated polyandry

### 2016-10-31 ms_pre_post.Rmd
Edit third and fourth paragraph of Introduction

@@

In a recent model, @Duthie showed that selection for alleles underlying polyandry and pre-copulatory inbreeding avoidance occurred only under highly constrained conditions in which direct costs of polyandry were low, and when very few males were available during a female's initial mate choice but many males were available during a female's additional mate choice(s), or when polyandry was conditionally expressed based on the degree to which a focal female engaged in inbreeding given her initially chosen mate. In the absence of these constraints, expression of polyandry was expected to increase inbreeding with each additional mate because remaining males were more likely to be relatives once available non-relatives were chosen. However, @Duthie assumed that polyandrous females had no control over the assignment of paternity to their offspring. If after engaging in polyandry, females could bias paternity through post-copulatory mechanisms, then selection for polyandry to avoid inbreeding might be more widely predicted when inbreeding depression is severe [@Zeh1997], potentially feeding back to drive further evolution of pre- and post-copulatory strategies.

>>>>

A recent genetically-explicit individual-based model showed that selection for alleles underlying polyandry and pre-copulatory inbreeding avoidance, and resulting joint evolution of these reproductive strategies, occurred only under highly constrained conditions [@Duthie]. Specifically, such evolution occurred when direct costs of polyandry were small, when very fewe males were available for a female's initial mate choice but many were available for additional mate choice(s), or when polyandry was conditionally expressed when a focal female was related to her initial mate [@Duthie]. In the absence of these conditions, increasing degrees of polyandry tended to increase the overall degree of inbreeding that females experienced, and hence increase inbreeding depression in offspring fitness. This occurred because polyandrous females that had evolved to avoid inbreeding had already chosen available unrelated males as their initial mates. Their additional mates, chosen from the remaining male population, were therefore increasingly likely to include relatives. Evolution of polyandry purely to facilitate pre-copulatory inbreeding avoidance was consequently restricted.

However, while @Duthie considered simultaneous evolution of polyandry and pre-copulatory inbreeding avoidance, they did not also allow simultaneous evolution of post-copulatory inbreeding avoidance, and assumed that polyandrous females had no post-copulatory control over fertilisation. If post-copulatory inbreeding avoidance could also evolve, polyandrous females could further bias paternity towards unrelated males within their set of chosen mates. This process might reduce the cost of polyandry stemming from accumulation of related mates, potentially facilitating evolution of polyandry to avoid inbreeding [@Zeh1997], and driving further consequent evolution of pre-copulatory or post-copulatory mate choice strategies. Therefore, while pre-copulatory inbreeding avoidance can facilitate evolution of polyandry under some restrictive conditions [@Duthie], the opportunity for post-copulatory choice might facilitate evolution of polyandry under a much wider range of conditions by allowing females to bias fertilisation toward non-kin. The long-term evolutionary dynamics of pre- and post-copulatory inbreeding avoidance then become unclear. Strong inbreeding depression might drive initial evolution of both pre- and post-copulatory inbreeding avoidance and associated polyandry. However, if evolution of polyandry and post-copulatory inbreeding avoidance renders pre-copulatory inbreeding avoidance redundant, or vice versa, then only one strategy might be maintained in the long-term.

@@
